Through the Eyes of the Teacher - Multimodal Exploration of Expertise Differences in the Perception of Classroom Disruptions in a Laboratory Study

### 2024-11-28 02_paper/02_study/02_expertise_paper/expertise_2024_11_18_MK.Rmd
Update expertise_2024_11_18_MK.Rmd

# Step 3: Extract only the participant's ID number (3 digits)
df_aoi <- df_aoi %>%
  mutate(Participant = as.numeric(str_extract(Participant, "\\d{3}")))

@@

```
### Check variables

@@

##################### AVERAGE DURATION ON AOI DISRUPTIVE PERSON ########################

@@

##################### NUMBER OF FIXATION ON AOI DISRUPTIVE PERSON ########################

# Define disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Filter and prepare data
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%
  dplyr::select(Group, Participant, TOI, Duration_of_interval, 
                "Number_of_fixations.Disruptive_Person") %>%
  mutate(
    Group = as_factor(Group),
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = Duration_of_interval / 1000
  ) %>%
  filter(!is.na(Disruption_Category))

# Summarize data
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarize(
    total_duration_sec = sum(Duration_of_interval_sec, na.rm = TRUE),
    fixation_count = sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(fixations_per_second = fixation_count / total_duration_sec)

# Individual-level aggregation
individual_summary <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarize(
    mean_fixations_per_second = mean(fixations_per_second, na.rm = TRUE),
    .groups = "drop"
  )

# 1. APA-Conform Table: Overall Fixation Rate
overall_stats <- individual_summary %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(mean_fixations_per_second, na.rm = TRUE), 2),
    SD = round(sd(mean_fixations_per_second, na.rm = TRUE), 2),
    Min = round(min(mean_fixations_per_second, na.rm = TRUE), 2),
    Max = round(max(mean_fixations_per_second, na.rm = TRUE), 2)
  )

# Render table for overall fixation rate
overall_table <- knitr::kable(
  overall_stats,
  caption = "Descriptive Statistics for Overall Fixation Rate (AOI Disruptive Person)",
  col.names = c("Group", "N", "M", "SD", "Min", "Max")
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Display overall table
overall_table

# 2. APA-Conform Table: Fixation Rate by Category Clustered by Disruption Category
category_stats <- disruption_summary %>%
  group_by(Disruption_Category, Group) %>%  # Group by category first, then by group
  summarise(
    N = n(),
    M = round(mean(fixations_per_second, na.rm = TRUE), 2),
    SD = round(sd(fixations_per_second, na.rm = TRUE), 2),
    Min = round(min(fixations_per_second, na.rm = TRUE), 2),
    Max = round(max(fixations_per_second, na.rm = TRUE), 2),
    .groups = "drop"  # Ungroup after summarization
  )

# Render the table clustered by Disruption Categories
category_table <- knitr::kable(
  category_stats,
  caption = "Descriptive Statistics for Fixation Rate by Disruption Category (AOI Disruptive Person)",
  col.names = c("Disruption Category", "Group", "N", "M", "SD", "Min", "Max")
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  kableExtra::group_rows(
    index = table(category_stats$Disruption_Category)  # Automatically cluster by category
  )

# Display the clustered category-specific table
category_table

# 3. Plot: Overall Fixation Rate
plot_overall <- individual_summary %>%
  ggplot(mapping = aes(x = Group, y = mean_fixations_per_second)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  ylim (0, 1) + 
  labs(x = "", y = "Mean Fixations per Second per Participant") +
  ggtitle("Overall Fixation Rate per Participant\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# 4. Plot: Fixation Rate by Disruption Category
plot_category <- disruption_summary %>%
  ggplot(mapping = aes(x = Group, y = fixations_per_second, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) + # Wrap long facet labels
  labs(x = "", y = "Fixations per Second") +
  ggtitle("Fixation Rate by Disruption Category\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10) # Adjust facet text size if necessary
  )

# Step 5: Statistical Tests by Category
category_results <- list()

for (category in unique(disruption_summary$Disruption_Category)) {
  cat_data <- disruption_summary %>% filter(Disruption_Category == category)
  t_cat <- t.test(fixations_per_second ~ Group, data = cat_data, var.equal = TRUE)
  d_cat <- cohen.d(fixations_per_second ~ Group, data = cat_data)
  category_results[[category]] <- data.frame(
    "Group Comparison" = "Experts vs. Novices",
    "t-value" = round(t_cat$statistic, 2),
    "df" = t_cat$parameter,
    "p-value" = round(t_cat$p.value, 3),
    "Mean Difference" = round(t_cat$estimate[1] - t_cat$estimate[2], 2),
    "Cohen's d" = round(d_cat$estimate, 2),
    "95% CI (d)" = paste0("[", round(d_cat$conf.int[1], 2), ", ", round(d_cat$conf.int[2], 2), "]")
  )
}

# Combine results into a single data frame and include Disruption Category
category_stats <- do.call(rbind, lapply(names(category_results), function(cat) {
  result <- category_results[[cat]]
  result$Disruption_Category <- cat  # Add Disruption Category as a column
  return(result)
}))

# Remove redundant Disruption Category column from the table
category_stats <- category_stats %>%
  relocate(Disruption_Category) %>%
  arrange(Disruption_Category) %>%  # Ensure proper order by category
  select(-Disruption_Category)  # Drop the column after adding grouping below

# Render the table with grouped rows
knitr::kable(
  category_stats,
  caption = "t-Test and Effect Size for Fixation Rate by Disruption Category",
  col.names = c("Group Comparison", "t-value", "df", "p-value", "Mean Difference", "Cohen's d", "95% CI (d)")
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  kableExtra::group_rows(
    group_label = "Lack of eagerness to learn",
    start_row = 1,
    end_row = 1
  ) %>%
  kableExtra::group_rows(
    group_label = "Physical disruptions",
    start_row = 2,
    end_row = 2
  ) %>%
  kableExtra::group_rows(
    group_label = "Verbal disruptions",
    start_row = 3,
    end_row = 3
  )

@@

# ET-Measures (AOI Disruptive Person)
# Number of fixations per second
# Define the different types of disruptions and categories
disruptions <- c(
  "Chatting_with_neighbour",
  "Clicking_pen",
  "Drawing",
  "Drumming_with_hands",
  "Head_on_table",
  "Heckling",
  "Looking_at_phone",
  "Snipping_with_fingers",
  "Whispering"
)
verbal_disruptions <- c("Chatting_with_neighbour",
                        "Heckling",
                        "Whispering")
physical_disruptions <- c("Clicking_pen",
                          "Drumming_with_hands",
                          "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone",
                       "Head_on_table",
                       "Drawing")

# Step 1: Filter and prepare the data
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%  # Filter rows based on the disruption list
  select(
    Group,
    Participant,
    TOI,
    Duration_of_interval,
    "Number_of_fixations.Disruptive_Person"
  ) %>%
  mutate(
    Group = as_factor(Group),  # Convert Group column to a factor
    Disruption_Category = case_when(
      # Assign categories based on disruption type
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2)  # Convert duration to seconds
  ) %>%
  filter(!is.na(Disruption_Category))  # Keep only rows with valid categories

# Step 2: Aggregate data by Participant and Disruption Category
df_aoi_numb_disrup <- df_aoi_disrup %>%
  group_by(Participant, Disruption_Category) %>%
  summarise(
    Total_Number_of_Fixations = sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE),
    Total_Duration_of_Interval_Sec = sum(Duration_of_interval_sec, na.rm = TRUE),
    .groups = "drop"  # Avoid unnecessary grouping
  )

# Step 3: Pivot the data to create a single row for each Participant
df_summary <- df_aoi_numb_disrup %>%
  pivot_wider(
    names_from = Disruption_Category,
    values_from = c(Total_Number_of_Fixations, Total_Duration_of_Interval_Sec),
    values_fill = 0  # Replace missing values with 0
  )

# SRI Rating Data
df_sri <- read_xlsx("data/Coding_SRI.xlsx") %>%
  filter(!ID %in% c(201, 223)) %>%
  rename(Participant = ID) %>%
  group_by(Participant) %>%
  summarise(
    Mean_disruption_appraisal = mean(disruption_appraisal, na.rm = TRUE),
    Mean_confidence_appraisal = mean(confidence_appraisal, na.rm = TRUE),
    Mean_prevalence_rating = mean(prevalence_rating, na.rm = TRUE)
  )

>>>>

# Step 3: Extract only the participant's ID number (3 digits) and assign group based on Participant ID
df_aoi <- df_aoi %>%
  mutate(
    Participant = as.numeric(str_extract(Participant, "\\d{3}")),
    Group = case_when(  # Assign group based on Participant ID
      Participant > 200 ~ "Expert",
      Participant < 200 ~ "Novice",
      TRUE ~ NA_character_  # In case there's any other value or missing
    )
  )

@@

### Check variables

@@

##################### AVERAGE DURATION ON AOI DISRUPTIVE PERSON ########################
##################### DATA PREPARATION ########################

@@

##################### TIME TO FIRST FIXATION ########################

# Define disruptions and categories
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Step 1: Data Preparation
df_ttff_disrup <- df_aoi %>%
  filter(TOI %in% c(verbal_disruptions, physical_disruptions, lack_of_eagerness)) %>%
  dplyr::select(
    Participant,
    TOI,
    Time_to_first_fixation.Disruptive_Person
  ) %>%
  mutate(
    Group = ifelse(Participant < 200, "Novice", "Expert"),  # Assigning Group based on Participant
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) %>% # Keep only rows with valid categories
  drop_na() %>%
  filter(
    Time_to_first_fixation.Disruptive_Person > 0, # Exclude zero fixation times
    Time_to_first_fixation.Disruptive_Person <= 30000 # Cut-off at 30 seconds
  ) %>%
  mutate(
    Disrup_time_fixation_sec = round(Time_to_first_fixation.Disruptive_Person / 1000, 2)
  )

# Step 2: Descriptive Statistics by Group
ttff_disrup_table <- df_ttff_disrup %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M in sec" = round(mean(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "SD in sec" = round(sd(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Min in sec" = round(min(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Max in sec" = round(max(Disrup_time_fixation_sec, na.rm = TRUE), 2)
  )

knitr::kable(ttff_disrup_table,
             caption = "N, M, SD, Min & Max Time to First Fixation in Seconds (AOI Disruptive Person)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### STATISTICS BY DISRUPTION CATEGORY ########################

# Step 3: Descriptive Statistics by Category and Group
summary_stats_category_group <- df_ttff_disrup %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    SD = round(sd(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    Median = round(median(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    IQR = round(IQR(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    Min = round(min(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    Max = round(max(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

knitr::kable(
  summary_stats_category_group,
  caption = "Descriptive Statistics for Time to First Fixation by Category and Group",
  col.names = c("Category", "Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Plot 1: Time to First Fixation by Group
plot_ttff_group <- ggplot(df_ttff_disrup, aes(x = Group, y = Disrup_time_fixation_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  labs(x = "Group", y = "Time to First Fixation (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Time to First Fixation by Group") +
  theme(legend.position = "none") +
  annotate("text", x = 1, y = max(df_ttff_disrup$Disrup_time_fixation_sec, na.rm = TRUE) * 1.1,
           label = paste0("Median = ", round(median(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"], na.rm = TRUE), 2),
                          "\nIQR = ", round(IQR(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"], na.rm = TRUE), 2)),
           hjust = 0.5, size = 4) +
  annotate("text", x = 2, y = max(df_ttff_disrup$Disrup_time_fixation_sec, na.rm = TRUE) * 1.1,
           label = paste0("Median = ", round(median(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"], na.rm = TRUE), 2),
                          "\nIQR = ", round(IQR(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"], na.rm = TRUE), 2)),
           hjust = 0.5, size = 4)

# Display the first plot
plot_ttff_group

# Plot 2: Time to First Fixation by Group and Disruption Category
plot_ttff_category <- ggplot(df_ttff_disrup, aes(x = Group, y = Disrup_time_fixation_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) +
  labs(x = "Group", y = "Time to First Fixation (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Time to First Fixation by Group and Disruption Category") +
  theme(legend.position = "none")

# Display the second plot
plot_ttff_category

# Normality Tests (Shapiro-Wilk)
normality_tests <- df_ttff_disrup %>%
  group_by(Group) %>%
  summarise(
    Shapiro_Wilk_p = round(shapiro.test(Disrup_time_fixation_sec)$p.value, 3),
    .groups = "drop"
  )

# Mann-Whitney U-Test and Effect Size
mann_whitney_test <- wilcox.test(Disrup_time_fixation_sec ~ Group, data = df_ttff_disrup, exact = FALSE)

# Effect Size Calculation for Mann-Whitney (r)
n1 <- nrow(df_ttff_disrup %>% filter(Group == "Novice"))
n2 <- nrow(df_ttff_disrup %>% filter(Group == "Expert"))
r_effect <- abs(mann_whitney_test$statistic - (n1 * n2 / 2)) / sqrt((n1 * n2 * (n1 + n2 + 1)) / 12)

kable(
  mann_whitney_results,
  caption = "Mann-Whitney U-Test Results for Time to First Fixation",
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Kruskal-Wallis Test and Pairwise Comparisons
kruskal_test <- kruskal.test(Disrup_time_fixation_sec ~ Disruption_Category, data = df_ttff_disrup)

# Pairwise Wilcoxon test with Bonferroni correction
pairwise_wilcox <- pairwise.wilcox.test(
  x = df_ttff_disrup$Disrup_time_fixation_sec,
  g = df_ttff_disrup$Disruption_Category,
  p.adjust.method = "bonferroni"
)

@@

# ET-Measures (AOI Disruptive Person)
# Number of fixations per second and Average duration of fixations
# Define the different types of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Step 1: Filter and prepare the data for both fixation counts and average durations
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%  # Filter rows based on the disruption list
  select(Group, Participant, TOI, 
         Duration_of_interval, 
         "Number_of_fixations.Disruptive_Person", 
         "Total_duration_of_fixations.Disruptive_Person") %>%
  mutate(
    Group = as_factor(Group),  # Convert Group column to a factor
    Disruption_Category = case_when(  # Assign categories based on disruption type
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),  # Convert duration to seconds
    Disrup_average_fixations = round(Total_duration_of_fixations.Disruptive_Person / 
                                     `Number_of_fixations.Disruptive_Person`, 2)  # Average duration in seconds
  ) %>%
  filter(!is.na(Disruption_Category))  # Remove rows with NA in disruption category

# Step 2: Summarize data by disruption category
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarise(
    total_duration_sec = round(sum(Duration_of_interval_sec, na.rm = TRUE), 2),  # Total time in seconds
    fixation_count = round(sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE), 2),  # Total fixations
    Disrup_fixations_per_second = round(fixation_count / total_duration_sec, 2),  # Fixation rate per second
    Disrup_average_fixations = round(mean(Disrup_average_fixations, na.rm = TRUE), 2),  # Average disruption fixations duration
    .groups = "drop"
  )

# Step 3: Summarize by participant
df_aoi_disrup <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarise(
    Disrup_fixations_per_second = round(mean(Disrup_fixations_per_second, na.rm = TRUE), 2),  # Mean fixation rate per participant
    Disrup_average_fixations = round(mean(Disrup_average_fixations, na.rm = TRUE), 2),  # Mean disruption fixation duration per participant
    .groups = "drop"
  ) %>% 
  select(Participant, Disrup_fixations_per_second, Disrup_average_fixations) %>%  # Select only relevant columns
  arrange(Participant)

# SRI Rating Data
df_sri <- read_xlsx("data/Coding_SRI.xlsx") %>%
  filter(!ID %in% c(201, 223)) %>%
  rename(Participant = ID) %>%
  group_by(Participant) %>%
  summarise(
    Disruption_appraisal = round(mean(disruption_appraisal, na.rm = TRUE), 2),
    Confidence_appraisal = round(mean(confidence_appraisal, na.rm = TRUE), 2),
    Prevalence_rating = round(mean(prevalence_rating, na.rm = TRUE), 2)
  ) %>% 
  arrange(Participant)
